Parallel Global Optimization for Non-Convex Mixed-Integer Problems \thanks{This study was supported by the Russian Science Foundation, project No.\,16-11-10150.}

### 2019-04-09 2019/RSD_Integer/RSD2019.tex
after translation

In the present paper, the mixed-integer global optimization problems are considered. A novel 
parallel algorithm for solving the problems of this class based on the information-statistical 
approach to solving the continuous global optimization problems has been proposed. 
The comparison of this algorithm with known analogs demonstrating the efficiency of the 
developed approach has been conducted. 
%Russian
Предложенный алгоритм допускает эффективное использование graphics accelerators. 
Проведены вычислительные эксперименты при решении серии из 100 multiextremal mixed-integer problems, которые подтверждают хорошее ускорение алгоритма с использованием GPU.

@@

In the present study, we proposed a novel parallel method for solving the mixed-integer 
problems based on the information-statistical approach to solving the global optimization 
problems \cite{Strongin2000,Strongin2013}. 
%Russian
В рамках данного подхода 
\begin{itemize}
	\item 
	решение многомерных задач сводится  к решению эквивалентных им одномерных, cоответствующая редукция основана на использовании space-filling curves;
	\item 
	при решении constrained optimization problems каждое ограничение учитывается и обрабатывается отдельно, штрафные функции не используются;
	\item 
	распараллеливание процесса поиска осуществляется за счет одновременного вычисления на каждой итерации нескольких значений целевой функции в разных точках search domain.	
\end{itemize}

Текст статьи, отражающей результаты проведенного исследования, построен следующим образом. 
В разделе 2 дано краткое описание описание схемы редукции размерности с использованием space-filling curves, а также описана индексная схема учета ограничений. Здесь же приведена формулировка параллельного индексного алгоритма для решения задач непрерывной глобальной оптимизации.
В разделе 3 предложен подход к обобщению параллельного индексного алгоритма с целью решения mixed-integer problems. Приведен способ, позволяющий свести решение mixed-integer problem к решению набора задач непрерывной оптимизации, которое может выполняться параллельно. 
Раздел 4 содержит результаты численных экспериментов. Здесь производится сравнение последовательной версии алгоритма с известными аналогами, а также показана эффективность параллельных CPU- и GPU-версий алгоритма при решении серии multiextremal mixed-integer problem.  
Section 5 concludes the paper.

@@

An efficient parallel index algorithm (PIA) for solving the constrained global optimization problem (\ref{problem}) has been developed at University of Nizhni Novgorod. The scheme of the algorithm is as follows.

@@

\begin{enumerate}
\item 
Renumber the points $x^1,...,x^k$ from previous iterations with lower indices, lowest to highest coordinate values, i.e.
\begin{equation}\label{Eq:17}
0=x_0<x_1<...<x_i<...<x_k<x_{k+1}=1,
\end{equation}
and match them with the values $z_i=g_\nu(y(x_i))$, $\nu=\nu(x_i)$, $1 \leq i \leq k$, from (\ref{trial_result}), calculated at these points; points $x_0=0$ и $x_{k+1}=1$ are introduced additionally; the values $z_0$ и $z_{k+1}$ are indeterminate.
\item
Classify the numbers $i,1\leq i \leq k$, of the trial points from (\ref{Eq:17}) by the number of problem constraints fulfilled at these points, by building the sets
\begin{equation}\label{Eq:18}
I_\nu = \left\{i: 1 \leq i \leq k,\ \nu = \nu(x_i)\right\},\ 1 \leq \nu \leq m+1,
\end{equation}
containing the numbers of all points $x_i,1\leq i \leq k$, with the same values of $\nu$. The end points $x_0=0$ and $x_{k+1}=1$ are interpreted as those with zero indices, and they are matched to an additional set $I_0={0,k+1}$.

Identify the maximum current value of the index
\begin{equation}\label{Eq:19}
M=\max \left\{\nu = \nu(x_i), \ 1\leq i \leq k\right\}.
\end{equation}
\item
For all values of $\nu, \ 1\leq \nu \leq m+1$, calculate the values  
\begin{equation}\label{Eq:20}
\mu_\nu = \max \left\{ \frac{\left|z_i-z_j\right|}{\left(x_i-x_j\right)^{1/N}} : i,j \in I_\nu, j<i\right\}.
\end{equation}
If the set $I_\nu$ contains less than two elements or $\mu_\nu$ from (\ref{Eq:20}) equals zero, then assume $\mu_\nu=1$.
\item
For all non-empty sets $I_\nu$, $1 \leq \nu \leq m+1$, determine the values
\begin{equation}\label{Eq:21}
  z^\ast_\nu =  
   \begin{cases}
    -\epsilon_\nu,  \nu < M, \\
    \min{\left\{g_\nu(x_i):i\in I_\nu\right\}}, \nu = M,
   \end{cases}
\end{equation}
where $M$ is the maximum current value of the index, and the vector
\begin{equation}\label{Eq:22}
\epsilon _R=\left(\epsilon_1,...,\epsilon_m\right),
\end{equation}
with positive coordinates is called the \textit{reserve vector} and is used as a parameter in the algorithm.
\item
For each interval $(x_{i-1},x_i)$,$1 \leq i \leq k+1$, calculate the \textit{characteristic} $R(i)$: 
\[
R(i)=\Delta_i+ \frac{(z_i-z_{i-1})^2}{(r_\nu\mu_\nu)^2\Delta_i}-2\frac{z_i+z_{i-1}-2z^\ast_\nu}{r_\nu\mu_\nu},\;\; \nu=\nu(x_{i-1})=\nu(x_i),
\]
\[
R(i)= 2\Delta_i-4\frac{z_i-z^\ast_\nu}{r_\nu\mu_\nu},\;\; \nu(x_{i-1})<\nu(x_i)=\nu,
\]
\[
R(i)= 2\Delta_i-4\frac{z_{i-1}-z^\ast_\nu}{r_\nu\mu_\nu},\;\; \nu = \nu(x_{i-1})>\nu(x_i).
\]
where $\Delta_i=(x_i-x_{i-1})^{1/N}$, and the values $r_\nu>1, 1\leq\nu\leq m+1$, are used as parameters in the algorithm.
\item
Reorder the characteristics $R(i)$, $1\leq i \leq k+1$, from highest to lowest 	
\begin{equation}\label{Eq:23}
R(t_1)\geq R(t_2)\geq ... \geq R(t_{k})\geq R(t_{k+1})
\end{equation}
and choose $p$ largest characteristics with interval numbers $t_j, 1\leq j \leq p$.
\item
Carry out $p$ new trials in parallel at the points $x^{k+j}, 1 \leq j \leq p$, calculated by the formulae
\begin{eqnarray*}
& x^{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2}, \; \nu(x_{t_j-1})\neq \nu(x_{t_j}), \\
& x^{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2}- \frac{\mathrm{sign}(z_{t_j}-z_{t_j-1})}{2r_\nu}\left[\frac{\left|z_{t_j}-z_{t_j-1}\right|}{\mu_\nu}\right]^N, \; \nu(x_{t_j-1})=\nu(x_{t_j})=\nu. \\
\end{eqnarray*}

@@

\section{Parallel algorithm for mixed-integer problems}
%Russian
Рассмотрим теперь случай, когда аргумент функций задачи содержит две компоненты: вектор $y$, принадлежащий гиперинтервалу $D$, и вектор $u$, имеющий конечный (и при этом не очень большой) набор возможных значений, т.е. 
\begin{gather}\label{problem_i}
\min{\left\{ g_{m+1}(y,u):y\in D, \; g_i(y,u)\leq 0, \; 1 \leq i \leq m\right\}},\\
D=\left\{a_j\leq y_j \leq b_j, \; 1\leq j \leq N \right\}.\nonumber
\end{gather}

Такие конечные наборы могут характеризовать, например, вариант материала, из которого создается объект, геометрические размеры или другие величины, которые могут принадлежать стандартному дискретному ряду, и т.п.

Занумеруем целыми числами $s, 1\leq s \leq S,$ все возможные значения вектора $u$, т.е. сопоставим каждому рассматриваемому значению $s$ вектор $u_s$. 
Тогда рассматриваемая задача может быть записана в виде 
\begin{gather}\label{problem_is}
 \min_{s\in\{1,...,S\}}\left(\min{\left\{ g_{m+1}(y,u_s):y\in D, \; g_i(y,u_s)\leq 0, \; 1 \leq i \leq m\right\}}\right),\\
D=\left\{ a_j\leq y_j \leq b_j, \; 1 \leq j\leq N \right\}.\nonumber 
\end{gather}

Используя схему редукции размерности с помощью развертки $y(x), x\in [0,1]$ можно сопоставить каждой вложенной задаче минимизации по $y$ одномерную задачу минимизации
\[
 \min{\left\{ g_{m+1}(y(x),u_s):x \in [0,1], \; g_i(y(x),u_s)\leq 0, \; 1 \leq i \leq m\right\}}, s\in\{1,...,S\}.
\]

Рассмотрим теперь соответствие 
\[
Y(x)=y(x-E(x)), \; x\in[0,S],
\]
отображающее любую точку интервала [0,S] на область $D$ (обозначение $E(x)$ соответствует целой части числа $x$) и определим функции 
\[
g_i(x) = g_i(Y(x),u_{E(x)+1}), x\in[0,S],
\]
имеющие, вообще говоря, разрывы первого рода в целочисленных точках $x_k = i, 1\leq i \leq S-1$.
Значение  $z_k = g_\nu(y(x_k))$ из (\ref{trial_result}) в этих точках будем считать неопределенным, а значения индекса -- равным 0, т.е. $\nu(x_k) = 0$ .

Используя введенные обозначения можно переформулировать исходную задачу как
\begin{equation}\label{problem_is1}
\min \left\{g_{m+1}(x): x \in [0,S], \; g_i(x) \leq 0, \; 1 \leq i \leq m\right\}.
\end{equation}

@@

Applying the parallel index algorithm to solving the problem (\ref{problem_is1}), we will find 
the solution of the problem (\ref{problem_i}). In this case the major part of trials will conducted in 
the subproblem, the solving of which corresponds to the solving of the initial 
problem (\ref{problem_i}). In the rest subproblems, a only minor part of trials will be performed 
since the solutions of these subproblems are the locally optimal ones.
%with respect to the solution of the $s$-th subproblem. 
All the above is confirmed by Fig. \ref{fig:1}, where the  
points of trials executed in the course of solving this problem are denoted by the dashes.

Thus, we have constructed the \textit{Mixed-Integer Parallel Index Algorithm} (MIPIA) based on the  
reduction of the mixed-integer non-convex optimization problem to the non-convex optimization problem. 
%Russian
Предложенная схема основана на параллельном индексном методе для решения задач непрерывной оптимизации и не является ориентированной на конкретное вычислительное устройство. Одной из ключевых операций здесь является параллельное проведение нескольких испытаний в различных точках области поиска (см. шаг 7 алгоритма), которе может быть реализовано как на CPU (с использованием OpenMP и/или MPI), так и на GPU (с использованием CUDA). Вопросы использования GPU для этих целей были детально рассмотрены в \cite{Barkalov2016,Gergel2016}, указанный подход к использованию GPU был применен и при реализации алгоритма MIPIA.

\section{Results of experiments}
%Russian
Первая серия экспериментов была проведена с использованием последовательной версии предложенного алгоритма с целью его сравнения с известными методами аналогичного назначения.

@@

Последующие серии экспериментов были выполнены для оценки ускорения параллельной версии предложенного алгоритма MIPIA с использованием как CPU, так и GPU. В данных экспериментах решалась серия из 100 тестовых mixed-integer problems, сгенерированных случайных образом.

Известным генератором тестовых задач непрерывной многоэкстремальной оптимизации является GKLS-генератор \cite{Gaviano}. Он позволяет генерировать functions of arbitrary dimensionality with known properties (the number of local minima, the size of their domains of attraction, the global minimizer, etc.).  
This generator of multiextremal functions is often used for the investigations of the global optimization algorithms~ 
\cite{Paulavicius2014,SergeyevKvasov2015,Lebedev2015,Gergel2015}. 
In Fig.~\ref{example} (a) and (b), the contour plots of two-dimensional GKLS functions are presented. Figures also shows the points of the trials performed by the method until the required accuracy $\epsilon=10^{-2}$ was achieved.

@@

В проведенных экспериментах GKLS-генератор использовался как основа для построения mixed-integer problems. Правила, позволяющие генерировать тестовые задачи подобного вида, заключаются в следующем.

\begin{enumerate}
	\item С помощью генератора GKLS порождается непрерывная многоэкстремальная функция $\varphi(y), \; y\in D, \; D = \left\{ a_j\leq y_j\leq b_j, 1\leq j \leq N \right\}$. Глобальный минимум данной функции достигается в известной точке $y'=(y'_1,...,y'_N)$ и равен $\varphi'=\varphi(y')=-1$.
	\item Генерируется concave mixed-integer function 
	\[
			\mu(y,u) = -2 \left[ \sum_{j=1}^N \left( \frac{y_j - y'_j}{b_j-a_j} \right)^2 + \sum_{j=1}^M \left( \frac{u_j - b_j}{b_j-a_j} \right)^2 \right],
	\]
	где 
	\begin{gather}
	y\in D = \left\{ a_j\leq y_j\leq b_j, 1\leq j \leq N \right\} \subset R^N,\nonumber \\
	u\in U = \left\{ u_j \in  \left\{a_j, ..., b_j \right\}, 1\leq j \leq M \right\};\nonumber
	\end{gather}
	т.е. данная функция имеет $N$ непрерывных и $M$ дискретных параметров и достигает минимального значения в точке $(y',b)$.
	\item Вычисляется коэффициент 
	\[
	C = 4 - \min_{y,u} \left\{ \mu(y,u) \right\}, \; y\in D, u \in U.
	\]
	Очевидно, что минимум вогнутой функции $\mu(y,u)$ в одной из угловых точек области поиска, поэтому при размерности задачи порядка 10 его вычисление может быть выполнено методом перебора.
	\item Формируется многоэкстремальная mixed-integer function 
	\[
	f(y,u) = \left(\varphi(y) + \sum_{j=1}^M{u_j}\right)\left(C + \mu(y,u)\right).
	\]
	По построению $f(y,u)$  будет принимать минимальное значение в точке $(y',b)$.

@@

Было сгенерировано 100 11-и мерных mixed-integer problems данного типа. The accuracy of the search was equal to $10^{-2}$.   
Computational experiments were carried out on Lobachevsky supercomputer. The node of 
supercomputer included two Intel Sandy Bridge E5-2660 2.2 GHz CPUs and 64 Gb RAM. The 
CPU had 8 cores, i.e. each node had a total of 16 cores and two NVIDIA Kepler K20Х GPUs.

В Table \ref{tab:2} приведены результаты экспериментов на CPU с использованием OpenMP в зависимости от числа задействованных threads $p$. Использовалось 1, 8 и 16 потоков. Приведено среднее число итераций $K_{av}$, требуемое для решения задачи, среднее время решения $T_{av}$ (в секундах), time speedup $S$ and iteration speedup $s$ (по отношению к последовательному запуску, т.е. $p=1$). В соответствии со схемой распараллеливания число испытаний на одной итерации параллельного алгоритма равнялось числу задействованных threads.

\begin{table}
	\caption{Результаты экспериментов на CPU}
	\label{tab:2}
	\center
	\begin{tabular}{cccccc}
		\hline\noalign{\smallskip}        
		$p$ & $T_{av}$ & $K_{av}$ & $S$ & $s$ \\
	\noalign{\smallskip} \hline \noalign{\smallskip}
	1 \;&	5520 \;  & 3221023 \;  & ...\; & ...  \\
	8 \;&	1253 \; & 512314 \;  & 4.4\; & 6.3  \\
	16\;&  717 \; &  209237 \; & 7.7\; & 15.4 \\
		\noalign{\smallskip}\hline
	\end{tabular}
\end{table}

Как видно из Table \ref{tab:2} , наблюдается почти линейное ускорение по итерациям и значительное ускорение по времени решения задачи, при этом последовательному алгоритму в среднем требовалось примерно 1.5 часа для решения задачи.

В Table \ref{tab:3} приведены результаты экспериментов на GPU с использованием CUDA в зависимости от числа задействованных GPU-threads. Приведено среднее число итераций $K_{av}$, требуемое для решения задачи, среднее время решения $T_{av}$ (в секундах), time speedup $S$ and iteration speedup $s$ (по отношению к полной загрузке CPU на узле кластера).

\begin{table}
	\caption{Результаты экспериментов на GPU}
	\label{tab:3}
	\center
	\begin{tabular}{cccccc}
		\hline\noalign{\smallskip}
			$p$ & $T_{av}$ & $K_{av}$ & $S$ & $s$ \\
	\noalign{\smallskip} \hline \noalign{\smallskip}
	256 \;&	33.6 \; & 1522\;  & 21.3 \; & 137  \\
	512 \;&	31.2 \; & 919 \;  & 23.0 \; & 228  \\
	1024\;& 30.2 \; & 412 \; & 23.7  \; & 508  \\
	2048\;& 38.5 \; & 244 \; & 18.6  \; & 858  \\
		\noalign{\smallskip}\hline
	\end{tabular}
\end{table}

Результаты эксперимнетов показывают почти одинаковое time speedup и линейное iteration speedup при использовании $p=256, 512, 1024$ threads. Однако при использовании $p=2048$ threads время работы алгоритма замедляется, но число итераций продолжает падать. Данный эффект объясняется тем, что для параллельного проведения $p$ испытаний используется GPU, а для обработки результатов испытаний (которая подразумевает работу со всей накопленной на предыдущих итерациях поисковой информацией) используется CPU. И при использовании $p=2048$ GPU threads время, которое тратится на передачу и обработку результатов 2048 испытаний, становится сопоставимым со временем проведения испытания, что и приводит к замедлению алгоритма в целом.

>>>>

In the present paper, the mixed-integer global optimization problems are considered. A novel 
parallel algorithm for solving the problems of this class based on the information-statistical 
approach to solving the continuous global optimization problems has been proposed. 
The comparison of this algorithm with known analogs demonstrating the efficiency of the 
developed approach has been conducted. 
%Russian
The proposed algorithm allows an efficient parallelization including the employment of the 
graphics accelerators. 
The results of performed numerical experiments (solving a series of 100 multiextremal mixed-
integer problems) confirm a good speed-up of the algorithm with the use of GPU.

@@

In the present study, we proposed a novel parallel method for solving the mixed-integer 
problems based on the information-statistical approach to solving the global optimization 
problems \cite{Strongin2000,Strongin2013}. 
%Russian
Within the framework of this approach: 
\begin{itemize}
	\item 
	the solving of the multidimensional problems is reduced to solving the equivalent one-
dimensional problems; the corresponding reduction is based on the use of the space-filling 
curves;
	\item 
	when solving the constrained optimization problems, each constraint is taken into 
account and processed separately, the penalty functions are not used;
	\item 
	the parallelization of the search process is performed by means of the simultaneous 
computing of several objective function values at different points of the search domain within 
every iteration.
\end{itemize}

The paper text reflecting the results of the performed study is organized in the following way. 
In Section 2, a brief description of the dimensionality reduction scheme using the space-filling 
curves is given. Also, the index scheme of accounting for the constraints is described. Here the 
formulation of the parallel index algorithm for solving the continuous global optimization 
problems is given as well.
In Section 3, the approach to the generalization of the parallel index algorithm with the purpose 
of solving the mixed-integer problems is presented. A method allowing reducing the solving of 
mixed-integer problem to solving a set of the continuous optimization problems, which can be 
performed in parallel is given. 
Section 4 contains the results of numerical experiments. The comparison of the sequential 
version of the algorithm with the known analogs is conducted here. Also, the efficiency of the 
parallel CPU- and GPU-versions of the algorithm for solving a series of the multiextremal 
mixed-integer problem is demonstrated. 
Section 5 concludes the paper.

@@

An efficient parallel index algorithm (PIA) for solving the constrained global optimization 
problem (\ref{problem}) has been developed at University of Nizhnii Novgorod. The scheme of 
the algorithm is as follows.

@@

\begin{enumerate}
\item 
Renumber the points $x^1,...,x^k$ from previous iterations with lower indices, lowest to highest 
coordinate values, i.e.
\begin{equation}\label{Eq:17}
0=x_0<x_1<...<x_i<...<x_k<x_{k+1}=1,
\end{equation}
and match them with the values $z_i=g_\nu(y(x_i))$, $\nu=\nu(x_i)$, $1 \leq i \leq k$, from 
(\ref{trial_result}), calculated at these points; points $x_0=0$ and $x_{k+1}=1$ are introduced 
additionally; the values $z_0$ and $z_{k+1}$ are indeterminate.
\item
Classify the numbers $i,1\leq i \leq k$, of the trial points from (\ref{Eq:17}) by the number of 
problem constraints fulfilled at these points, by building the sets
\begin{equation}\label{Eq:18}
I_\nu = \left\{i: 1 \leq i \leq k,\ \nu = \nu(x_i)\right\},\ 1 \leq \nu \leq m+1,
\end{equation}
containing the numbers of all points $x_i,1\leq i \leq k$, with the same values of $\nu$. The end 
points $x_0=0$ and $x_{k+1}=1$ are interpreted as those with zero indices, and they are 
matched to an additional set $I_0={0,k+1}$.

Identify the maximum current value of the index
\begin{equation}\label{Eq:19}
M=\max \left\{\nu = \nu(x_i), \ 1\leq i \leq k\right\}.
\end{equation}
\item
For all values of $\nu, \ 1\leq \nu \leq m+1$, calculate the values  
\begin{equation}\label{Eq:20}
\mu_\nu = \max \left\{ \frac{\left|z_i-z_j\right|}{\left(x_i-x_j\right)^{1/N}} : i,j \in I_\nu, 
j<i\right\}.
\end{equation}
If the set $I_\nu$ contains less than two elements or $\mu_\nu$ from (\ref{Eq:20}) equals zero, 
then assume $\mu_\nu=1$.
\item
For all non-empty sets $I_\nu$, $1 \leq \nu \leq m+1$, determine the values
\begin{equation}\label{Eq:21}
  z^\ast_\nu =  
   \begin{cases}
    -\epsilon_\nu,  \nu < M, \\
    \min{\left\{g_\nu(x_i):i\in I_\nu\right\}}, \nu = M,
   \end{cases}
\end{equation}
where $M$ is the maximum current value of the index, and the vector
\begin{equation}\label{Eq:22}
\epsilon _R=\left(\epsilon_1,...,\epsilon_m\right),
\end{equation}
with positive coordinates is called the \textit{reserve vector} and is used as a parameter in the 
algorithm.
\item
For each interval $(x_{i-1},x_i)$,$1 \leq i \leq k+1$, calculate the \textit{characteristic} $R(i)$: 
\[
R(i)=\Delta_i+ \frac{(z_i-z_{i-1})^2}{(r_\nu\mu_\nu)^2\Delta_i}-2\frac{z_i+z_{i-1}-
2z^\ast_\nu}{r_\nu\mu_\nu},\;\; \nu=\nu(x_{i-1})=\nu(x_i),
\]
\[
R(i)= 2\Delta_i-4\frac{z_i-z^\ast_\nu}{r_\nu\mu_\nu},\;\; \nu(x_{i-1})<\nu(x_i)=\nu,
\]
\[
R(i)= 2\Delta_i-4\frac{z_{i-1}-z^\ast_\nu}{r_\nu\mu_\nu},\;\; \nu = \nu(x_{i-1})>\nu(x_i).
\]
where $\Delta_i=(x_i-x_{i-1})^{1/N}$, and the values $r_\nu>1, 1\leq\nu\leq m+1$, are used as 
parameters in the algorithm.
\item
Reorder the characteristics $R(i)$, $1\leq i \leq k+1$, from highest to lowest 	
\begin{equation}\label{Eq:23}
R(t_1)\geq R(t_2)\geq ... \geq R(t_{k})\geq R(t_{k+1})
\end{equation}
and choose $p$ largest characteristics with interval numbers $t_j, 1\leq j \leq p$.
\item
Carry out $p$ new trials in parallel at the points $x^{k+j}, 1 \leq j \leq p$, calculated by the 
formulae
\begin{eqnarray*}
& x^{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2}, \; \nu(x_{t_j-1})\neq \nu(x_{t_j}), \\
& x^{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2}- \frac{\mathrm{sign}(z_{t_j}-z_{t_j-
1})}{2r_\nu}\left[\frac{\left|z_{t_j}-z_{t_j-1}\right|}{\mu_\nu}\right]^N, \; \nu(x_{t_j-
1})=\nu(x_{t_j})=\nu. \\
\end{eqnarray*}

@@

\section{Parallel algorithm for mixed-integer problems}
%Russian
Now let us consider the case when the argument of the problem functions consists of two 
components: the vector $y$ falling into the hyperinterval $D$ and the vector $u$ taking a finite 
(and not too large) set of possible values i. e. 
\begin{gather}\label{problem_i}
\min{\left\{ g_{m+1}(y,u):y\in D, \; g_i(y,u)\leq 0, \; 1 \leq i \leq m\right\}},\\
D=\left\{a_j\leq y_j \leq b_j, \; 1\leq j \leq N \right\}.\nonumber
\end{gather}

Such finite sets can characterize, for example, the variant of material, which the object is made 
from, the geometric sizes, or other quantities, which can belong to a standard discrete series, etc.

Let us number by the integer values $s, 1\leq s \leq S,$ all possible values of the vector $u$ i. e. 
superimpose to each considered value $s$ to the vector $u_s$. 
Then, the considered problem can be written in the form 
\begin{gather}\label{problem_is}
 \min_{s\in\{1,...,S\}}\left(\min{\left\{ g_{m+1}(y,u_s):y\in D, \; g_i(y,u_s)\leq 0, \; 1 \leq i \leq 
m\right\}}\right),\\
D=\left\{ a_j\leq y_j \leq b_j, \; 1 \leq j\leq N \right\}.\nonumber 
\end{gather}

Using the dimensionality reduction scheme with the evolvent $y(x), x\in [0,1]$, one can 
superimpose each nested minimization problem with respect to $y$ to a one-dimensional 
problem 
\[
 \min{\left\{ g_{m+1}(y(x),u_s):x \in [0,1], \; g_i(y(x),u_s)\leq 0, \; 1 \leq i \leq m\right\}}, 
s\in\{1,...,S\}.
\]

Now let us consider the relation 
\[
Y(x)=y(x-E(x)), \; x\in[0,S],
\]
mapping any point of the interval [0,S] onto the domain $D$ (the notation $E(x)$ corresponds 
to the integer part of the number $x$) and define the functions 
\[
g_i(x) = g_i(Y(x),u_{E(x)+1}), x\in[0,S],
\]
having, in general, the discontinuities of the first kind at the integer points $x_k = i, 1\leq i \leq 
S-1$.
The values $z_k = g_\nu(y(x_k))$ from (\ref{trial_result}) at these points will be considered to 
be undefined, and the values of the indices -- to equal to 0 i. e. $\nu(x_k) = 0$.

Using the introduced notations, one can reformulate the original problem as
\begin{equation}\label{problem_is1}
\min \left\{g_{m+1}(x): x \in [0,S], \; g_i(x) \leq 0, \; 1 \leq i \leq m\right\}.
\end{equation}

@@

Applying the parallel index algorithm to solving the problem (\ref{problem_is1}), we will find 
the solution of the problem (\ref{problem_i}). In this case the major part of trials will conducted 
in 
the subproblem, the solving of which corresponds to the solving of the initial 
problem (\ref{problem_i}). In the rest subproblems, only the minor part of trials will be 
performed 
since the solutions of these subproblems are the locally optimal ones.
%with respect to the solution of the $s$-th subproblem. 
All the above is confirmed by Fig. \ref{fig:1}, where the  
points of trials executed in the course of solving this problem are denoted by the dashes.

Thus, we have constructed the \textit{Mixed-Integer Parallel Index Algorithm} (MIPIA) based 
on the  
reduction of the mixed-integer non-convex optimization problem to the non-convex 
optimization problem. 
%Russian
The proposed computational scheme is based on the parallel index method for solving the 
continuous optimization problems and is not oriented onto any particular computational device. 
Here the parallel performing of several trials at different points of the search domain is one of 
key operations (see Step 7 of the algorithm), which can be implemented on CPU (with the use 
of OpenMP and/or MPI) as well as on GPU (with the use CUDA). The issues of the use of 
GPU for these purposes have been considered in details in 
Ref.~\cite{Barkalov2016,Gergel2016}. This approach to the use of GPU has been applied in 
the implementation of the MIPIA algorithm as well.

\section{Results of experiments}
%Russian
The first series of experiments was conducted using the sequential version of the proposed 
algorithm in order to compare this one with well known methods of similar purpose.

@@

The next series of experiments were performed in order to evaluate the speedup of the parallel 
version of the proposed MIPIA algorithm with the use of CPU as well as GPU. In these 
experiments, a series of 100 mixed-integer test problems generated in a random way was solved.

GKLS is a well known generator of the test problems for the continuous multiextremal 
optimization~\cite{Gaviano}. It allows generating the functions of arbitrary dimensionality with 
known properties (the number of local minima, the size of their domains of attraction, the global 
minimizer, etc.).  
This generator of multiextremal functions is used for the investigations of the global 
optimization algorithms 
often~\cite{Paulavicius2014,SergeyevKvasov2015,Lebedev2015,Gergel2015}. 
In Fig.~\ref{example} (a) and (b), the contour plots of two-dimensional GKLS functions are 
presented. Figures also shows the points of the trials performed by the method until the required 
accuracy $\epsilon=10^{-2}$ was achieved.

@@

In the performed experiments, the GKLS generator was used as a base for the construction the 
mixed-integer problems. The rules, allowing generating the test problems of this kind consist in 
the following.

\begin{enumerate}
	\item A continuous multiextremal function $\varphi(y), \; y\in D, \; D = \left\{ a_j\leq 
y_j\leq b_j, 1\leq j \leq N \right\}$ is generated with the use of the GKLS generator. The global 
minimum of this function is achieved at the known point $y'=(y'_1,...,y'_N)$ and equals 
$\varphi'=\varphi(y')=-1$.
	\item A concave mixed-integer function 
	\[
			\mu(y,u) = -2 \left[ \sum_{j=1}^N \left( \frac{y_j - y'_j}{b_j-a_j} 
\right)^2 + \sum_{j=1}^M \left( \frac{u_j - b_j}{b_j-a_j} \right)^2 \right],
	\]
	is generated where 
	\begin{gather}
	y\in D = \left\{ a_j\leq y_j\leq b_j, 1\leq j \leq N \right\} \subset R^N,\nonumber \\
	u\in U = \left\{ u_j \in  \left\{a_j, ..., b_j \right\}, 1\leq j \leq M \right\};\nonumber
	\end{gather}
	i. e. this function has $N$ continuous and $M$ discrete parameters and achieves its 
minimum value at the point $(y',b)$.
	\item The coefficient 
	\[
	C = 4 - \min_{y,u} \left\{ \mu(y,u) \right\}, \; y\in D, u \in U.
	\]
	is computed. 
It is obvious, that the minimum of a concave function $\mu(y,u)$ is located at one of the corner 
points of the search domain. Therefore, if the problem dimensionality is of the order of 10 it can 
be found by the brute force method.
	\item A multiextremal mixed-integer function 
	\[
	f(y,u) = \left(\varphi(y) + \sum_{j=1}^M{u_j}\right)\left(C + \mu(y,u)\right).
	\]
	is formed.
By construction, $f(y,u)$ would take its minimum value at the point $(y',b)$.

@@

Total 100 11-dimensional mixed-integer problems of this kind were generated. The accuracy of 
the search was equal to $10^{-2}$.   
Computational experiments were carried out on Lobachevsky supercomputer. The node of 
supercomputer included two Intel Sandy Bridge E5-2660 2.2 GHz CPUs and 64 Gb RAM. The 
CPU had 8 cores, i.e. each node had a total of 16 cores and two NVIDIA Kepler K20X GPUs.

In Table \ref{tab:2}, the results of experiments on CPU with the use of OpenMP are presented 
subject to the number of employed threads $p$. Total 1, 8, and 16 threads were used. The 
average number of iterations $K_{av}$ required to solve the problem, the average time of 
solving $T_{av}$ (in seconds), the time speedup $S$, and iteration speedup $s$ (with respect 
to the sequential run i. e. for $p=1$) are presented. In accordance with the parallelization 
scheme, the number of trials within a single iteration of the parallel algorithm was equal to the 
number of employed threads.

\begin{table}
	\caption{The results of experiments on CPU}
	\label{tab:2}
	\center
	\begin{tabular}{cccccc}
		\hline\noalign{\smallskip}        
		$p$ & $T_{av}$ & $K_{av}$ & $S$ & $s$ \\
	\noalign{\smallskip} \hline \noalign{\smallskip}
	1 \;&	5520 \;  & 3221023 \;  & ...\; & ...  \\
	8 \;&	1253 \; & 512314 \;  & 4.4\; & 6.3  \\
	16\;&  717 \; &  209237 \; & 7.7\; & 15.4 \\
		\noalign{\smallskip}\hline
	\end{tabular}
\end{table}

As one can see from Table \ref{tab:2}, almost linear speedup in iterations and an essential 
speedup in the problem solving time have been observed. At that, the sequential algorithm spent 
almost 1.5 hours in average to solve a problem.

In Table \ref{tab:3}, the results of experiments on GPU with the use of CUDA are presented 
subject to the number of GPU threads employed. The average number of iterations $K_{av}$ 
required to solve a problem, the average solving time $T_{av}$ (in seconds), the time speedup 
$S$, and iteration speedup $s$ (with respect to the full load of CPU on a cluster node) are 
presented.

\begin{table}
	\caption{The results of experiments on GPU}
	\label{tab:3}
	\center
	\begin{tabular}{cccccc}
		\hline\noalign{\smallskip}
			$p$ & $T_{av}$ & $K_{av}$ & $S$ & $s$ \\
	\noalign{\smallskip} \hline \noalign{\smallskip}
	256 \;&	33.6 \; & 1522\;  & 21.3 \; & 137  \\
	512 \;&	31.2 \; & 919 \;  & 23.0 \; & 228  \\
	1024\;& 30.2 \; & 412 \; & 23.7  \; & 508  \\
	2048\;& 38.5 \; & 244 \; & 18.6  \; & 858  \\
		\noalign{\smallskip}\hline
	\end{tabular}
\end{table}

The results of experiments demonstrate almost the same time speedup and linear iteration 
speedup when using $p=256, 512,$ and $1024$ threads. However, when using $p=2048$ 
threads, the algorithm worktime increased but the number of iterations continued to decrease. 
This effect was explained by the fact that for parallel running of $p$ trials GPU is used but  for 
processing the results of the trials (which implies the processing of the whole search information 
accumulated during the preceding iterations) CPU is uses. And when employing $p=2048$ 
GPU threads, the time, which is spent for the transfer and processing of the results of 2048 
trials became comparable to the time of executing the trials that leads to the slowing down of 
the algorithm as a whole.

\section{Conclusion}
%Russian
In the present paper, the results obtained in Lobachevskii State University of Nizhnii Novgorod 
when developing and investigating the parallel global optimization algorithms for solving the 
multiextremal problems, in which some parameters are continuous and the others are the discrete 
ones are presented. 
An efficient parallel algorithm has been proposed for solving the problems of the specified class. 
In the sequential variant, this algorithm does not inferior to similar purpose algorithm 
implemented in Matlab Global Optimization Toolbox.
In the parallel variant, the proposed algorithm allows the efficient implementations on CPU as 
well as on GPU. 
The numerical experiments
on a series of 100 multiextremal mixed-integer test problems have been carried out 
demonstrating a good speedup of the algorithm with the use of GPU convincingly.
Thus, the parallel algorithm with the use of GPU demonstrated the time speedup $S = 182$ 
relative to the sequential one and $S = 23.7$ relative to the algorithm employing two CPU on 
the cluster node fully.
